Kids health and child data cleaning

### 2020-06-22 child_health_data_cleaning.Rmd
data cleaning updates

---
title: "Kids health and child data cleaning"
author: "Ethan Walker"
date: "Started 27 Feb 2020, Updated 28 May 2020"
output: pdf_document 
---

@@

```{r}
# Load data
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))

demographics_clean <- 
  read_rds(paste0(file_path, "Output/demographics_clean.rds"))

@@

# Load and merge files from different locations
```{r}
wmt_athomelog_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_athomelog_20200227.xlsx")) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "WMT",
         at_home_day = as.factor(Day),
         # rename columns and recode as factor
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

nn_athomelog_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_athomelog_20200416.xlsx")) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "NN",
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

ak_athomelog_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_athomelog_20200227.xlsx")) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "AK",
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

# Combine logs from all areas
kids_at_home_log <- rbind(wmt_athomelog_initial, nn_athomelog_initial,
                          ak_athomelog_initial) %>% 
  left_join(kids_linked_ids, c("area", "home_winter_id"))

@@

  select(area, home, home_winter_id, child_id_num, winter_id, treatment,
         sampling_day, sample_time, at_home, "00:00:00":"23:00:00") %>% 

@@

  select(area, home, home_winter_id, child_id_num, winter_id, treatment,
         sampling_day, sample_time, at_home, at_home_location)  %>% 

@@

# start with cleaned PM dataset
at_home_pm_merged <- pm_clean %>% 
  group_by(area, home) %>% 
  # average the PM for sampling period, day, and hour
  mutate(pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_sampling_day = mean(pm, na.rm = TRUE),
         pm_datetime_merge = round_date(pm_datetime_new, unit = "hour"),
         pm_obs_sampling_day = n()) %>% 
  separate(pm_datetime_merge, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  group_by(area, home, sampling_day, sample_time) %>% 
  mutate(pm_mean_hour = mean(pm, na.rm = TRUE)) %>% 
  arrange(area, home_winter_id, sampling_day, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  # keep one observation per hour of sampling, with PM averaged
  distinct(sample_time, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(sampling_day = as.character(sampling_day)) %>% 
  select(area, home, home_winter_id, winter_id, treatment, sampling_day, 
         sample_date, sample_time, pm_mean_hour, pm_mean_sampling_day, 
         pm_mean_sampling_period, day_of_week, pm_obs_sampling_day, 
         pm_spike_1000, pm_spike_2000, pm_spike_5000) %>% 
  # join with at home log from above
  left_join(kids_at_home_log_new, 
            by = c("area", "home", "home_winter_id", "winter_id", "treatment",
                   "sampling_day", "sample_time")) %>% 
  group_by(area, home, home_winter_id, child_id_num, at_home) %>% 
  # average PM during time child was at home
  mutate(pm_at_home_sampling_period = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  group_by(area, home, home_winter_id, child_id_num, sampling_day, at_home) %>% 
  mutate(pm_at_home_sampling_day = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  arrange(area, home_winter_id, child_id_num, sampling_day, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  distinct(at_home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  filter(!is.na(at_home)) %>% 
  arrange(area, home_winter_id, child_id_num, at_home, sampling_day)

@@

# Load and merge files from different locations
```{r}
wmt_medicalvisit_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_medicalvisit_20200227.xlsx")) %>% 
  # recode values to NA
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>%
  # recode temps of <90 to NA
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  select(home_winter_id:med_visit_mean_temp)

nn_medicalvisit_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_medicalvisit_20200416.xlsx")) %>%
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "NN",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>% 
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  replace_with_na(replace = list(med_visit_mean_temp = "NaN")) %>%
  select(home_winter_id:med_visit_mean_temp)

ak_medicalvisit_initial <- read_xlsx(paste0(file_path, "Input/AK/ak_medicalvisit_20200416.xlsx")) %>% 
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "AK",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.character(ARI),
         med_visit_lethargy = as.character(Lethargy),
         med_visit_retractions = as.character(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.character(Wheeze),
         med_visit_stridor = as.character(Stridor),
         med_visit_crackles = as.character(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.character(Referred),
         med_visit_notes = as.character(Level)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>% 
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>%
  replace_with_na(replace = list(med_visit_mean_temp = "NaN")) %>%
  select(home_winter_id:med_visit_mean_temp)

# Combine logs from all areas
kids_med_visit <- rbind(wmt_medicalvisit_initial, nn_medicalvisit_initial,
                          ak_medicalvisit_initial) %>% 
  left_join(kids_linked_ids, c("area", "home_winter_id"))

@@

# Load and merge files from different locations
```{r}
wmt_birthinfancy_initial <- read_xlsx(paste0(file_path, "Input/WMT/wmt_birthinfancy_20200227.xlsx")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         gender = as.character(Gender),
         bw_lbs = as.numeric(BirthweightLbs),
         bw_oz = as.numeric(BirthWeightOz),
         birth_weight_oz = as.numeric((bw_lbs*16)+bw_oz),
         mom_age = as.numeric(AgeOfMom),
         breastfeed_current = as.character(CurrentlyBreastfeeding),
         breastfeed_past = as.character(DidBreastfeed),
         bf_years = as.numeric(HowLongYears),
         bf_months = as.numeric(HowLongMonths),
         breastfeed_months = as.numeric((bf_years*12)+bf_months),
         flu_shot = as.character(FluShot),
         flu_shot_planned = as.character(FluShotPlanned),
         asthma_meds = as.character(AsthmaMeds),
         child_school = as.character(Attend),
         school_hours = as.numeric(HoursDay),
         school_days = as.numeric(DaysWeek),
         school_hours_week = as.numeric(school_hours*school_days),
         sibling_school = as.character(SibSchool)) %>% 
  replace_with_na(replace = list(bw_lbs = 0, birth_weight_oz = 0, mom_age = 0)) %>%
  select(home_winter_id:sibling_school)

nn_birthinfancy_initial <- read_xlsx(paste0(file_path, "Input/NN/nn_birthinfancy_20200416.xlsx")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "NN",
         gender = as.character(Gender),
         bw_lbs = as.numeric(BirthweightLbs),
         bw_oz = as.numeric(BirthWeightOz),
         birth_weight_oz = as.numeric((bw_lbs*16)+bw_oz),
         mom_age = as.numeric(AgeOfMom),
         breastfeed_current = as.character(CurrentlyBreastfeeding),
         breastfeed_past = as.character(DidBreastfeed),
         bf_years = as.numeric(HowLongYears),
         bf_months = as.numeric(HowLongMonths),
         breastfeed_months = as.numeric((bf_years*12)+bf_months),
         flu_shot = as.character(FluShot),
         flu_shot_planned = as.character(FluShotPlanned),
         asthma_meds = as.character(AsthmaMeds),
         child_school = as.character(Attend),
         school_hours = as.numeric(HoursDay),
         school_days = as.numeric(DaysWeek),
         school_hours_week = as.numeric(school_hours*school_days),
         sibling_school = as.character(SibSchool)) %>% 
  replace_with_na(replace = list(bw_lbs = 0, birth_weight_oz = 0, mom_age = 0)) %>%
  select(home_winter_id:sibling_school)

>>>>

---
title: "Kids health and child data cleaning"
author: "Ethan Walker"
date: "Started 27 Feb 2020, Updated 22 June 2020"
output: pdf_document 
---

@@

```{r}
# Load data
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 
demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds"))

@@

# Load and merge files from different locations
```{r}
wmt_athomelog_initial <- read_csv(paste0(file_path, "Input/WMT/athomelog.csv"), 
                                  na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "WMT",
         at_home_day = as.factor(Day),
         # rename columns and recode as factor
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

nn_athomelog_initial <- read_csv(paste0(file_path, "Input/NN/athomelog.csv"), 
                                  na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "NN",
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

ak_athomelog_initial <- read_csv(paste0(file_path, "Input/AK/athomelog.csv"), 
                                  na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         child_id_num = as.factor(ChildID),
         area = "AK",
         at_home_day = as.factor(Day),
         at_home_7 = as.factor(AtHome1),
         location_7 = as.factor(Where1),
         at_home_8 = as.factor(AtHome2),
         location_8 = as.factor(Where2),
         at_home_9 = as.factor(AtHome3),
         location_9 = as.factor(Where3),
         at_home_10 = as.factor(AtHome4),
         location_10 = as.factor(Where4),
         at_home_11 = as.factor(AtHome5),
         location_11 = as.factor(Where5),
         at_home_12 = as.factor(AtHome6),
         location_12 = as.factor(Where6),
         at_home_13 = as.factor(AtHome7),
         location_13 = as.factor(Where7),
         at_home_14 = as.factor(AtHome8),
         location_14 = as.factor(Where8),
         at_home_15 = as.factor(AtHome9),
         location_15 = as.factor(Where9),
         at_home_16 = as.factor(AtHome10),
         location_16 = as.factor(Where10),
         at_home_17 = as.factor(AtHome11),
         location_17 = as.factor(Where11),
         at_home_18 = as.factor(AtHome12),
         location_18 = as.factor(Where12),
         at_home_19 = as.factor(AtHome13),
         location_19 = as.factor(Where13),
         at_home_20 = as.factor(AtHome14),
         location_20 = as.factor(Where14),
         at_home_21 = as.factor(AtHome15),
         location_21 = as.factor(Where15),
         at_home_22 = as.factor(AtHome16),
         location_22 = as.factor(Where16)) %>% 
  select(home_winter_id:location_22)

# Combine logs from all areas
kids_at_home_log <- rbind(wmt_athomelog_initial, nn_athomelog_initial,
                          ak_athomelog_initial) %>% 
  mutate(area = as.factor(area)) %>% 
  right_join(kids_linked_ids, c("area", "home_winter_id")) %>% 
  filter(!is.na(child_id_num))

@@

  select(area, home, home_winter_id, child_id_num, winter_id, treatment_assigned,
         treatment_actual, sampling_day, sample_time, at_home, "00:00:00":"23:00:00") %>% 

@@

  select(area, home, home_winter_id, child_id_num, winter_id, treatment_assigned,
         treatment_actual, sampling_day, sample_time, at_home, at_home_location)  %>% 

@@

# start with cleaned PM dataset
at_home_pm_merged <- pm_clean %>% 
  group_by(area, home) %>% 
  # average the PM for sampling period, day, and hour
  mutate(pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_sampling_day = mean(pm, na.rm = TRUE),
         pm_datetime_merge = round_date(pm_datetime_new, unit = "hour"),
         pm_obs_sampling_day = n()) %>% 
  separate(pm_datetime_merge, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  group_by(area, home, sampling_day, sample_time) %>% 
  mutate(pm_mean_hour = mean(pm, na.rm = TRUE)) %>% 
  arrange(area, home_winter_id, sampling_day, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  # keep one observation per hour of sampling, with PM averaged
  distinct(sample_time, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(sampling_day = as.character(sampling_day)) %>% 
  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
         sampling_day, sample_date, sample_time, pm_mean_hour, pm_mean_sampling_day, 
         pm_mean_sampling_period, day_of_week, pm_obs_sampling_day, 
         pm_spike_1000, pm_spike_2000, pm_spike_5000) %>% 
  # join with at home log from above
  left_join(kids_at_home_log_new, 
            by = c("area", "home", "home_winter_id", "winter_id", "treatment_assigned",
                   "treatment_actual", "sampling_day", "sample_time")) %>% 
  group_by(area, home, home_winter_id, child_id_num, at_home) %>% 
  # average PM during time child was at home
  mutate(pm_at_home_sampling_period = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  group_by(area, home, home_winter_id, child_id_num, sampling_day, at_home) %>% 
  mutate(pm_at_home_sampling_day = mean(pm_mean_hour, na.rm = TRUE)) %>% 
  arrange(area, home_winter_id, child_id_num, sampling_day, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  distinct(at_home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  filter(!is.na(at_home)) %>% 
  arrange(area, home_winter_id, child_id_num, at_home, sampling_day)

@@

# Load and merge files from different locations
```{r}
wmt_medicalvisit_initial <- read_csv(paste0(file_path, "Input/WMT/medicalvisit.csv"), 
                                  na = c("NULL", "")) %>% 
  # recode values to NA
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         # fix data entry error for date
         med_visit_date = as.character(VisitDate),
         med_visit_date = if_else(med_visit_date == "2025-11-04", "2015-11-04", med_visit_date),
         med_visit_date = ymd(med_visit_date),
         med_visit_ari = as.factor(ARI),
         med_visit_lethargy = as.factor(Lethargy),
         med_visit_retractions = as.factor(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.factor(Wheeze),
         med_visit_stridor = as.factor(Stridor),
         med_visit_crackles = as.factor(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.factor(Referred)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>%
  # recode temps of <90 to NA
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread back to wide format
  spread(sample, value) %>%
  replace_with_na(replace = list(med_visit_mean_temp = "NaN")) %>%
  select(home_winter_id:med_visit_mean_temp)

nn_medicalvisit_initial <- read_csv(paste0(file_path, "Input/NN/medicalvisit.csv"), 
                                  na = c("NULL", "")) %>%
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "NN",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.factor(ARI),
         med_visit_lethargy = as.factor(Lethargy),
         med_visit_retractions = as.factor(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.factor(Wheeze),
         med_visit_stridor = as.factor(Stridor),
         med_visit_crackles = as.factor(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.factor(Referred)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>% 
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread back to wide format
  spread(sample, value) %>%
  replace_with_na(replace = list(med_visit_mean_temp = "NaN")) %>%
  select(home_winter_id:med_visit_mean_temp)

ak_medicalvisit_initial <- read_csv(paste0(file_path, "Input/AK/medicalvisit.csv"), 
                                  na = c("NULL", "")) %>% 
  mutate(HRPulseOx = if_else(HRPulseOx < 50, -999, HRPulseOx),
         HR = if_else(HR < 50, -999, HR),
         SP02 = if_else(SP02 < 70, -999, SP02)) %>% 
  replace_with_na_all(condition = ~.x %in% c(-999, -9999, 0)) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "AK",
         med_visit_date = ymd(VisitDate),
         med_visit_ari = as.factor(ARI),
         med_visit_lethargy = as.factor(Lethargy),
         med_visit_retractions = as.factor(Retractions),
         temp1 = as.numeric(Temp1),
         temp2 = as.numeric(Temp2),
         temp3 = as.numeric(Temp3),
         med_visit_resp_rate = as.numeric(RR),
         med_visit_heart_rate = as.numeric(HR),
         med_visit_wheeze = as.factor(Wheeze),
         med_visit_stridor = as.factor(Stridor),
         med_visit_crackles = as.factor(Crackles),
         med_visit_sp02 = as.numeric(SP02),
         med_visit_hr_sp02 = as.numeric(HRPulseOx),
         med_visit_referred = as.factor(Referred)) %>% 
  # gather 3 temperatures to calculate average
  gather("sample", "value", c("temp1":"temp3")) %>% 
  group_by(home_winter_id, child_id_num, med_visit_date) %>% 
  mutate(value = if_else(value < 90, -99, value)) %>% 
  replace_with_na(replace = list(value = -99)) %>% 
  mutate(med_visit_mean_temp = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread back to wide format
  spread(sample, value) %>%
  replace_with_na(replace = list(med_visit_mean_temp = "NaN")) %>%
  select(home_winter_id:med_visit_mean_temp)

# Combine logs from all areas
kids_med_visit <- rbind(wmt_medicalvisit_initial, nn_medicalvisit_initial,
                          ak_medicalvisit_initial) %>% 
  mutate(area = as.factor(area)) %>% 
  right_join(kids_linked_ids, c("area", "home_winter_id"))

summary(kids_med_visit)

@@

# Load and merge files from different locations
```{r}
wmt_birthinfancy_initial <- read_csv(paste0(file_path, "Input/WMT/birthinfancy.csv"), 
                                  na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "WMT",
         gender = as.factor(Gender),
         bw_lbs = as.numeric(BirthweightLbs),
         bw_oz = as.numeric(BirthWeightOz),
         birth_weight_oz = as.numeric((bw_lbs*16)+bw_oz),
         mom_age = as.numeric(AgeOfMom),
         breastfeed_current = as.factor(CurrentlyBreastfeeding),
         breastfeed_past = as.factor(DidBreastfeed),
         bf_years = as.numeric(HowLongYears),
         bf_months = as.numeric(HowLongMonths),
         breastfeed_months = as.numeric((bf_years*12)+bf_months),
         flu_shot = as.factor(FluShot),
         flu_shot_planned = as.factor(FluShotPlanned),
         asthma_meds = as.factor(AsthmaMeds),
         child_school = as.factor(Attend),
         school_hours = as.numeric(HoursDay),
         school_days = as.numeric(DaysWeek),
         school_hours_week = as.numeric(school_hours*school_days),
         sibling_school = as.factor(SibSchool)) %>% 
  replace_with_na(replace = list(bw_lbs = 0, birth_weight_oz = 0, mom_age = 0)) %>%
  select(home_winter_id:sibling_school)

nn_birthinfancy_initial <- read_csv(paste0(file_path, "Input/NN/birthinfancy.csv"), 
                                  na = c("NULL", "")) %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         child_id_num = as.character(ChildID),
         area = "NN",
         gender = as.factor(Gender),
         bw_lbs = as.numeric(BirthweightLbs),
         bw_oz = as.numeric(BirthWeightOz),
         birth_weight_oz = as.numeric((bw_lbs*16)+bw_oz),
         mom_age = as.numeric(AgeOfMom),
         breastfeed_current = as.factor(CurrentlyBreastfeeding),
         breastfeed_past = as.factor(DidBreastfeed),
         bf_years = as.numeric(HowLongYears),
         bf_months = as.numeric(HowLongMonths),
         breastfeed_months = as.numeric((bf_years*12)+bf_months),
         flu_shot = as.factor(FluShot),
         flu_shot_planned = as.factor(FluShotPlanned),
         asthma_meds = as.factor(AsthmaMeds),
         child_school = as.factor(Attend),
         school_hours = as.numeric(HoursDay),
         school_days = as.numeric(DaysWeek),
         school_hours_week = as.numeric(school_hours*school_days),
         sibling_school = as.factor(SibSchool)) %>% 
  replace_with_na(replace = list(bw_lbs = 0, birth_weight_oz = 0, mom_age = 0)) %>%
  select(home_winter_id:sibling_school)
